\section{Evaluation and Limitations}

The project reached its main goal of booting a custom kernel on the Raspberry Pi4 and combining task scheduling, interrupt handling, UART control, HDMI output, Sense HAT support, and runtime diagnostics in one system. The result is usable for demonstration and experimentation, especially because tasks can be started and stopped interactively and the system can show its internal state through the shell and HDMI dashboards.

However, several parts of the implementation remain fragile. The I2C subsystem is the most important example. Although access to the bus is serialized with a global lock, communication with the Sense HAT devices is still timing-sensitive and occasionally unreliable. This affects joystick input, sensor reads, and LED matrix updates, because all of them share the same physical bus. The locking mechanism prevents obvious concurrent transactions, but it does not fully eliminate lower-level communication problems such as missed device responses, incomplete transfers, or recovery after a failed transaction.

The HDMI implementation also required several iterations. A simple line-based redraw approach was too expensive and interfered with the cooperative scheduler. The final dirty-cell model improved this by flushing only changed character cells, but it also made the display code more complex. Correct pane ownership, cursor handling, scrolling, and partial redraws are therefore still more error-prone than simple serial output over UART.

The main architectural limitation is the cooperative scheduler. It is compact and understandable, but it depends on every task yielding, sleeping, or blocking regularly. A task that performs long-running work without yielding can reduce system responsiveness. In addition, all code runs in one privileged address space, so there is no user/kernel separation, virtual memory, process isolation, filesystem support, or memory protection.

Future work should therefore focus less on adding more demo tasks and more on robustness. The I2C layer would benefit from better error handling, timeouts, bus recovery, and clearer transaction-state tracking. A preemptive scheduler would make the system more responsive under faulty or CPU-heavy tasks. Stronger memory and stack checks would make task execution safer, and persistent storage could be added for logs, configuration, or collected sensor data.